\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-05-07 paper/manuscript_paNote.tex
got started on pipeline section, not done yet

Once we have gathered the website files we need to pre-process the data \citep{denny2018text}. In this section, we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text files that are suitable for comparative analysis with text as data methods. In this pipeline, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

\subsection{Site to Text Conversion}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For all text analysis projects, researchers need to consider file formats. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}.

Consequently, we rename all documents so that their file ending reflects their actual file type. This is strictly necessary because we rely on the \texttt{readtext} R package \citep{readtext}, which determines a document's type solely through its ending---to convert the files to plain text.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

@@

The last round of preprocessing is intended to remove everything from the file that is not an English word. This step is tailored to our intention to use the text for topic modeling. The final preprocessing round includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive\footnote{Topic models essentially do not pick up on extremely rare words, so their inclusion is a waste of computational resources. Removing them in this manner is also the default preprocessing choice in the \texttt{stm} package.} to our models while increasing their computational cost, we also discard any token types that occur in only one document. We also conduct lemmatization to reduce words to their basic form. Lemmatization is similar to stemming but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible. To this end, we rely on the R package \texttt{spacyr}, which provides an R implementation of the Python library \texttt{spaCy}.

\subsection{Boilerplate Removal}\label{boilerplate}
As noted above, city websites contain a large amount of text that is uninformative for its actual content, and therefore a hindrance to understanding through algorithmic text analysis. This is a common issue with textual data in which informative content is embedded in technically structured documents. See, e.g., \citet{burgess2016legislative,wilkerson2015tracing} and \citet{linder2018text} for examples of boilerplate removal in the analysis of legislative text. In the case of websites, lines in documents are generally quite informative, so all of our boilerplate removal efforts are done at the line level.

%this is outdated; not deleting it yet in case we need it for further reference
%Consequently we remove this content as follows: Each line of every document is compared to every line in every other document belonging to the same city. We count how many times each line is duplicated for that city. We remove any line occurring more than our chosen threshold of 10.\footnote{Empirically, lines tend to be duplicated either hundreds of times or only once or twice, if at all.} This means that each document only retains the information that is particular about it. We implement this algorithm through hash tables, which reduces the computational complexity from O($N^2$) to O($N$). Before this step is taken, we remove numbers and dates from the documents because they frequently make lines unique, despite the fact that they are virtually the same (for example different days on a city calendar).

\subsection*{Boilerplate Classification}
In order to determine whether a line should be discarded, we train a classifier on a human-coded sample. We sampled 500 lines from documents in each of the following five cities: Los Angeles, CA, Indianapolis, IN, New York, NY, Shreveport, LA, and Seattle, WA. To ensure that lines which occur more frequently in these cities (sometimes hundreds of thousands of times) had a higher probability of being scrutinized by the classifier, we use sampling weights equivalent to the proportion of total lines in a city's corpus made up by each specific line type. As an example, the most common line throughout all pages of the city of Seattle consists only of the word ``total'' and occurs 103,068 times. Similarly, the line ``page'' occurs 58,833 times. Even something completely nonsensical such as ``a a'' still appears on 376 occasions. To account for the higher likelihood of some lines being part of the training set, we use inverse probability weights in training the classifier---the weight of each line in the sample is 1/[number of occurrences in the corpus].\footnote{Note that the performance of the classifier is robust to the use of these weights and only changes by about one percentage point if they are not used.}

These 500 lines were then hand-coded as either substantively informative (210 lines) or not (290 lines). We then trained a number of different classifiers with this informativeness measure as the dependent variable. The independent variables we use are: (1) number of times the line was duplicated within the city, (2) the length of the line, in characters, (3) the number of tokens in the line, and (4) the median distance from the document midpoint to the position of the line itself. The purpose of these covariates is as follows:

\begin{itemize}
\item {\bf Line length:} The length of the line and the number of tokens are ways to find lines consisting of only a word or two. This is highly predictive of lines which are used as website headers and navigational elements, which are of zero substantive interest to us but are very effective at differentiating cities. These terms also happen to be fairly common, which causes them to be overweighted by the topic model.

\item {\bf Number of line duplications:} To directly address the latter problem, we include a measure of the number of times a line is duplicated within a city. Many lines occur hundreds or even thousands of times on a single website and therefore are terms that are highly predictive of the website, which causes the topic model to find topics that are highly predictive of cities, but not substantively informative.

\item {\bf Line position in the document} Since boilerplate terms such as navigational elements, headers, footers, and so on, should occur more frequently at the beginning and the end of websites, we attempt to identify such content as following: We measure the distance between the midpoint of a document and the position of a line, expressed as quantiles (to account for differing document lengths). Since lines can occur in multiple documents or multiple times in the same document, we take the median of these measures. Thus, for example, a line which often occurs at the beginning of documents might have a score of 0.45, whereas a line that tends to be found more in the center, and thus is indicative of more relevant content, might be scored with a 0.11 instead.
\end{itemize}

%Finally, the keyness measure: This indicator shows whether a term occurs disproportionally frequently in one document collection, compared to another. In our case, this would be one city's documents, compared to all others. Mathematically, this is a simple chi-square test, asserting whether the frequency of a word in a city is greater than its expected count, as determined by its count in all other documents, and the number of tokens in the city. The variable used in the model is the chi-square value of this test for each word (which varies and is therefore calculated individually for each city).\footnote{Or alternatively, the log of the absolute chi-square value, after which the previously negative values are made negative again.}

%A regression table for this model can be found in table 3. The associated accuracy in 5-fold cross-validation is 0.8325. The cross-validation accuracy of a simplified model with only the number of characters as a covariate is 0.825 (table 4). A model with the four covariates described above as well as interaction terms between each of them yields an accuracy of 0.885 (table 5).

%\input{tables/boilerplateClassifier1.tex}
\input{tables/boilerplateClassifierRFMetrics.tex}
%\input{tables/boilerplateClassifierRFImportance.tex}

We rely on a random forest as the final classifier, which offers slightly better performance than logistic regression.\footnote{We also tried SVM, boosted trees and AdaBoost, with similar results and chose the random forests because this method has a probabilistic basis and is more intuitive.} We assess the performance of this classifier through five-fold cross-validation. This means that the classifier is trained on 400 samples and then tested on the held-out set of 100, measuring metrics such as percent correctly predicted, precision, recall, and F1 score. This procedure is carried out five times so that each sample is part of the test set once. The aggregated (mean) results of this process can be found in Table \ref{randomForest}. For the implementation of this method, we rely on the R package \texttt{caret}, whose random forest classifier is based on the package \texttt{ranger}. We use this classifier to flag and remove all lines that are not classified (based on a threshold of $p = 0.5$) as substantively meaningful. The effect of this process on the corpus is illustrated with the corpus of Anchorage, AK (i.e. a city that isn't part of our sample used in the analysis) as an example in Figures \ref{boilerplate_before_after1} to \ref{boilerplate_before_after4}. Before the lines identified by the classifier as boilerplate are removed, lines with very few characters and words are the most common. After the removal, the distribution has changed---lines of medium length now occur more frequently than extremely short ones, which are unlikely to be substantively meaningful (see figures \ref{boilerplate_before_after1} and \ref{boilerplate_before_after2}). Furthermore, lines that are duplicated only a few times rather than dozens, hundreds or even thousands are now more common (see figure \ref{boilerplate_before_after3}). Finally, the position of the line in the documents is not as important to the random forest, and this also shows in the results. However, this feature still has a positive effect, as lines at either end of the document are a bit less common now (see figure \ref{boilerplate_before_after4}). Table \ref{tabBoilerplateIllustration} provides further illustration by listing the top 10 most likely boilerplate lines (in Anchorage, AK) -- all of which were flagged as such with a probability of 1. After all the preprocessing is set and done, our corpus consists of 259,099 documents.

\begin{figure}[htp]
    \centering
    \caption{Effects of the boilerplate classifier on the corpus of the city of Anchorage, AK. After the boilerplate content is removed, extremely short lines are less common.}
    \label{boilerplate_before_after1}
    \includegraphics[width=\linewidth]{figures/boilerplateBeforeAfterNwords.pdf}
\end{figure}
\begin{figure}[htp]
    \centering
    \caption{Effects of the boilerplate classifier on the corpus of the city of Anchorage, AK. After the boilerplate content is removed, extremely short lines are less common.}
    \label{boilerplate_before_after2}
    \includegraphics[width=\linewidth]{figures/boilerplateBeforeAfterNchars.pdf}
\end{figure}
\begin{figure}[htp]
    \centering
    \caption{Effects of the boilerplate classifier on the corpus of the city of Anchorage, AK. After the boilerplate content is removed, lines that are duplicated hundreds or thousands of times are less common.}
    \label{boilerplate_before_after3}
    \includegraphics[width=\linewidth]{figures/boilerplateBeforeAfterFreq.pdf}
\end{figure}
\begin{figure}[htp]
    \centering
    \caption{Effects of the boilerplate classifier on the corpus of the city of Anchorage, AK. After the boilerplate content is removed, lines at the beginning and end of documents are less common.}
    \label{boilerplate_before_after4}
    \includegraphics[width=\linewidth]{figures/boilerplateBeforeAfterMedianDocMidDist.pdf}
\end{figure}

>>>>

Once we have gathered the website files we need to pre-process the data \citep{denny2018text}. In this section, we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents that are suitable for comparative analysis with text as data methods. In this pipeline, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

\subsection{Site to Text Conversion}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For all text analysis projects, researchers need to consider file formats. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}. Files with incorrect endings are subsequently renamed.

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus and navigational elements. We leverage methods developed in the information retrieval literature to deal with this problem. These so-called 'boilerplate detection' tools rely on both structural features, such as HTML tags, as well as text statistics such as word and sentence length to estimate whether a given portion of an HTML file is useful. We rely on the method described in \cite{Kohlschuetter2010}, which has been published in a reputable conference, is widely cited and can easily be implemented through the R package \texttt{boilerpipeR}.

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package\footnote{\texttt{readtext} determines a document's type solely through its ending -- so the conversion described above is strictly necessary for the package to work correctly.} \citep{readtext}, which is a wrapper for a set of parsers.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.}\footnote{\texttt{readtext} also contains an HTML parser, but this tool simply extracts all text and is therefore inferior to boilerpipe.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

For preprocessing, we rely on a set of tools which are used much more commonly in natural langauge processing than political methodology -- part-of-speech (POS) tagging and named entity recognition (NER), as these procedures perform very well in separating substantively interesting information from any remaining boilerplate. Names of city officials and citizen petitioners feature frequently in city documents. The same is true for streets, locations and not least of all, the city itself. Since individual names recur at a much higher rate within a city than across the entire corpus, this causes a topic model to cluster its topics by city. As these names convey no substantive information, we rely on NER to remove them. Specifically we use \texttt{spacyR} to detect and remove all entities except X, Y and Z, as well. Similarly we remove words on the basis of their POS-tags, retaining only nouns (by far the most informative category), verbs, adjectives. Furthermore, we keep proper nouns that also occur as nouns -- this removes names, but retains titles such as ``Police Chief'' which can appear as proper nouns if they are followed by a name.

We also conduct lemmatization to reduce words to their basic form. Lemmatization is similar to stemming but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible.

Furthermore, we remove words with less than three characters (these are usually artifacts from improperly encoded documents and faulty or impartial optical character recognition), stopwords and non-English words (using the R package \texttt{Hunspell})\footnote{As we are using \texttt{spacyR} with an English tagger, non-English words are erroneously categorized as (proper) nouns.}.

A final and crucial step is the removal of duplicate documents, which occur very frequently on websites. In addition to their primary purpose, the previous preprocessing steps also help in stripping otherwise identical documents of information that makes them unique -- such as names and dates -- thus facilitating their deletion. The removal of duplicate information also solves a problem posed by a recurring feature of municipal websites -- city calendars. For most days in most cities, these calendars tend to be empty, but nevertheless contain a template, which would lead to the creation of a document in our corpus. These short and repetitive texts would pose severe problems to a topic model. Our approach preserves the valuable information contained within these calendars when cities do make use of them, but removes them when they offer no substantive data.

@@

The last round of preprocessing is intended to remove everything from the file that is not an English word. This step is tailored to our intention to use the text for topic modeling. The final preprocessing round includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive\footnote{Topic models essentially do not pick up on extremely rare words, so their inclusion is a waste of computational resources. Removing them in this manner is also the default preprocessing choice in the \texttt{stm} package.} to our models while increasing their computational cost, we also discard any token types that occur in only one document.

\subsection{Boilerplate Removal}\label{boilerplate}
As noted above, city websites contain a large amount of text that is uninformative for its actual content, and therefore a hindrance to understanding through algorithmic text analysis. This is a common issue with textual data in which informative content is embedded in technically structured documents. See, e.g., \citet{burgess2016legislative,wilkerson2015tracing} and \citet{linder2018text} for examples of boilerplate removal in the analysis of legislative text.

%\subsection*{Boilerplate Classification}
%In order to determine whether a line should be discarded, we train a classifier on a human-coded sample. We sampled 500 lines from documents in each of the following five cities: Los Angeles, CA, Indianapolis, IN, New York, NY, Shreveport, LA, and Seattle, WA. To ensure that lines which occur more frequently in these cities (sometimes hundreds of thousands of times) had a higher probability of being scrutinized by the classifier, we use sampling weights equivalent to the proportion of total lines in a city's corpus made up by each specific line type. As an example, the most common line throughout all pages of the city of Seattle consists only of the word ``total'' and occurs 103,068 times. Similarly, the line ``page'' occurs 58,833 times. Even something completely nonsensical such as ``a a'' still appears on 376 occasions. To account for the higher likelihood of some lines being part of the training set, we use inverse probability weights in training the classifier---the weight of each line in the sample is 1/[number of occurrences in the corpus].\footnote{Note that the performance of the classifier is robust to the use of these weights and only changes by about one percentage point if they are not used.}

%These 500 lines were then hand-coded as either substantively informative (210 lines) or not (290 lines). We then trained a number of different classifiers with this informativeness measure as the dependent variable. The independent variables we use are: (1) number of times the line was duplicated within the city, (2) the length of the line, in characters, (3) the number of tokens in the line, and (4) the median distance from the document midpoint to the position of the line itself. The purpose of these covariates is as follows:

After all the preprocessing is set and done, our corpus consists of 259,099 documents.
